Optimizing Abstract$^{\text 2}$ Machines

### 2012-11-11 paper/paper.tex
Removed fixpoint specialization step.

\begin{align*}
(cs, \msto) &\widehat{\machstep} (cs', \msto') \\
\mbox{where } CS' &= lift(\machstep)(\{ wn(c, \msto) \mid c \in cs\}) \\
              cs' &= \{nw(C') \mid C' \in CS'\} \\
              \msto' &= \bigcup\{C'.\msto \mid C' \in CS'\}
\end{align*}

@@

\begin{itemize}
 \item{[SS] quotient by sets of values (lazy non-determinism)}
 \item{[SS] remove {\tt ev} states (abstract compilation)}
 \item{[DS] specialize the fixpoint operator to work with widened
   machine configurations (fixpoint specialization)}
 \item{[DS] operate on store changes rather than whole stores (store deltas)}
 \item{[DS] represent the store as a global, mutable hash and
   determine a state's store by the number of changes made to the
   global store at the time of visiting (timestamp)}
 \item{[DS] use a global mutable workset}
 \item{[DS] represent the global store as a global, mutable vector in
   the 0-CFA case. (preallocation)}
\end{itemize}
Each technique requires minimal changes to the reduction
relation/fixpoint, and each improves performance by integer
factors. The first four remain in the purely functional realm. The
final three use mutation in small, localized places, so they do not
increase conceptual complexity drastically.

@@

\newpage
\subsection{Fixed-point specialization}

\begin{alltt}
;; \(\widehat{State} \to \widehat{State}\)
(define (wide-step-specialized state)
  (match state
    [(cons \(\msto\) cs)
     (define-values (cs* \(\msto\)*)
       (for/fold ([cs* \(\varnothing\)] [\(\msto\)* \(\msto\)])
         ([c cs])
         (match (step-compiled^ (cons \(\msto\) c))
           [(cons \(\msto\)** cs**)
            (values (\(\cup\) cs* cs**) (\(\sqcup\)-store \(\msto\)* \(\msto\)**))])))
     (cons \(\msto\)* (\(\cup\) cs cs*))]))
\end{alltt}
%% Should be commutted to beginning

Special + Compile + Lazy:
   cpu time: 14212 real time: 14681 gc time: 823

>>>>

\begin{align*}
(cs, \msto) &\widehat{\machstep} (cs', \msto') \\
\mbox{where } cs' &= \{c' \mid wn(c, \msto) \machstep wn(c, \msto^c), c \in cs\} \\
              \msto' &= \bigcup\{\msto^c \mid wn(c,\msto)\machstep wn(c', \msto^c), c \in cs\}
\end{align*}

@@

\begin{itemize}
 \item{[SS] quotient by sets of values (lazy non-determinism)}
 \item{[SS] remove {\tt ev} states (abstract compilation)}
 \item{[DS] operate on store changes rather than whole stores (store deltas)}
 \item{[DS] represent the store as a global, mutable hash and
   determine a state's store by the number of changes made to the
   global store at the time of visiting (timestamp)}
 \item{[DS] use a global mutable workset}
 \item{[DS] represent the global store as a global, mutable vector in
   the 0-CFA case. (preallocation)}
\end{itemize}
Each technique requires minimal changes to the reduction
relation/fixpoint, and each improves performance by integer
factors. The first three remain in the purely functional realm. The
final three use mutation in small, localized places, so they do not
increase conceptual complexity drastically.

@@

\newpage
